---
title: "Assessment 3: Group Project"
team: "Team Echo"
session: "Hexa 2, 2024"
coursecode: "ZZSC9020"
authors:
  - "Chuang, Keith – z5449930"
  - "Gandhi, Rupesh – z5368767"
  - "Melbin, Joseph – z5394849"
  - "CHIU, Yiu Tong – z5039191"
date: "2024-04-20"

Acknowledgements: 
  - "TBC."
  - "TBC."
  - "TBC."
  - "TBC."

output: html_document

---

# Abstract

---

# Contents

---

# 1.0 Introduction

---

# 2.0 Literature Review

## 2.1 Industry/Operational Context

### 2.1.1 Variables of Interest
#### 2.1.1.1 Temperature
#### 2.1.1.2 Population
#### 2.1.1.3 Industry Base
#### 2.1.1.4 Renewable Engery Prevalence

## 2.2 Research Topic

## 2.3 Statistical Models

### 2.3.1 Regression Analysis

#### 2.3.1.1 Linear Models 

#### 2.3.1.2 Generalised Models

#### 2.3.1.3 Logistics Models

#### 2.3.1.4 Poisson Models

#### 2.3.1.5 Log-Linear Models

#### 2.3.1.6 Model Assessment

#### 2.3.1.7 Polynomial Models

#### 2.3.1.8 Splines

#### 2.3.1.9 Generalised Additive Models

### 2.3.2 Multivariate Analysis

---

# 3.0 Material and Methods

# Packages

```{r}

library(readr)
library(here)
library(ggplot2)

```

---

# 4.0 Exploratory Data Analysis

## 4.1 Area of Research
Determine the efficacy of temperature as the independent variable in prediction 
of energy demand in the subsequent 24 hours on a half hourly basis for NSW 
utilizing statistical models i.e. regression. Explore the predictive improvement 
through the addition of independent variables. Do these models provide 
comparable predictive results to the supplied predictions and if so under what 
circumstances.

## 4.2 Collecting and Loading Data

### 4.2.1 Provided Data 

``` {r}

# Load Data

# VIC
data_forecastdemand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/a/forecastdemand_vic.csv"))

data_temperature_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_vic.csv"))

data_demand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_vic.csv"))

# SA
data_forecastdemand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/b/forecastdemand_sa.csv"))

data_temperature_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_sa.csv"))

data_demand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_sa.csv"))

# QLD
data_forecastdemand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/c/forecastdemand_qld.csv"))

data_temperature_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_qld.csv"))

data_demand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_qld.csv"))

# NSW
data_forecastdemand_nsw <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/NSW/forecastdemand_nsw.csv"))

data_temperature_nsw <-  read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/NSW/temperature_nsw.csv"))

data_demand_nsw <-  read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/NSW/totaldemand_nsw.csv"))

```

### 4.2.2 Sourced Data



## 4.3 Initial Data Inspection

``` {r}
# Generate Some Useful Functions

# check for null values
check_for_na <- function(data) {
  for (col_index in seq_along(data)) {
    has_na <- any(is.na(data[[col_index]]))
    cat("Column", col_index, "has NA values:", has_na, "\n")
  }
}

# check for unique values
#check_unique_entries <- function(data) {
#  unique_entries_list <- list()
#  for (col_name in names(data)) {
#    unique_entries <- unique(data[[col_name]])
#    unique_entries_list[[col_name]] <- unique_entries
#  }
#  return(unique_entries_list)
#}

```

``` {r}
# VIC
# check structure
str(data_forecastdemand_vic)
str(data_temperature_vic)
str(data_demand_vic)

# check for null values
check_for_na(data_forecastdemand_vic)
check_for_na(data_temperature_vic)
check_for_na(data_demand_vic)

# check character data types for uniqueness
unique(data_forecastdemand_vic$REGIONID)
unique(data_temperature_vic$LOCATION)
unique(data_demand_vic$REGIONID) # ensure REGIONID matches
```

``` {r}
# SA
# check structure
str(data_forecastdemand_sa)
str(data_temperature_sa)
str(data_demand_sa)

# check for null values
check_for_na(data_forecastdemand_sa)
check_for_na(data_temperature_sa)
check_for_na(data_demand_sa)

# check character data types for uniqueness
unique(data_forecastdemand_sa$REGIONID)
unique(data_temperature_sa$LOCATION)
unique(data_demand_sa$REGIONID) # ensure REGIONID matches
```

``` {r}
# QLD
# check structure
str(data_forecastdemand_qld)
str(data_temperature_qld)
str(data_demand_qld)

# check for null values
check_for_na(data_forecastdemand_qld)
check_for_na(data_temperature_qld)
check_for_na(data_demand_qld)

# check character data types for uniqueness
unique(data_forecastdemand_qld$REGIONID)
unique(data_temperature_qld$LOCATION)
unique(data_demand_qld$REGIONID) # ensure REGIONID matches
```

``` {r}
# NSW
# check structure
str(data_forecastdemand_nsw)
str(data_temperature_nsw)
str(data_demand_nsw)

# check for null values
check_for_na(data_forecastdemand_nsw)
check_for_na(data_temperature_nsw)
check_for_na(data_demand_nsw)

# check character data types for uniqueness
unique(data_forecastdemand_nsw$REGIONID)
unique(data_temperature_nsw$LOCATION)
unique(data_demand_nsw$REGIONID) # ensure REGIONID matches
```

## 4.4 Summary Statistics

``` {r}
# detect outliers, based on z score
detect_outliers <- function(data) {
  z_scores <- scale(data)
  outliers <- abs(z_scores) > 3
  return(outliers)
}

# detect outliers, based on modified z score
detect_modified_outliers <- function(data) {
  median_val <- median(data)
  mad_val <- mad(data)
  modified_z_scores <- (data - median_val) / mad_val
  outliers_modified <- abs(modified_z_scores) > 3.5
  return(outliers_modified)
}

```

``` {r}
# VIC

summary(data_forecastdemand_vic)
summary(data_temperature_vic)
summary(data_demand_vic)

boxplot(data_forecastdemand_vic$FORECASTDEMAND)
boxplot(data_temperature_vic$TEMPERATURE)
boxplot(data_demand_vic$TOTALDEMAND)

hist(data_forecastdemand_vic$FORECASTDEMAND)
hist(data_temperature_vic$TEMPERATURE)
hist(data_demand_vic$TOTALDEMAND)

#data_forecastdemand_vic_z_scores <- scale(data_forecastdemand_vic$FORECASTDEMAND)
#data_forecastdemand_vic_outliers <- abs(data_forecastdemand_vic_z_scores) > 3
#data_forecastdemand_vic_outliers_true <- data_forecastdemand_vic_outliers[data_forecastdemand_vic_outliers == TRUE]
#print(data_forecastdemand_vic_outliers == TRUE)

detect_outliers(data_forecastdemand_vic$FORECASTDEMAND)

```

``` {r}



```

``` {r}
```

``` {r}
```

## 4.5 Data Visualisation

``` {r} 
# QLD Data
ggplot(data_demand_qld, aes(x = DATETIME, y = TOTALDEMAND)) +
  geom_point() +
  labs(x = "X-axis label", y = "Y-axis label", title = "Scatter Plot")
```

## 4.6 Data Relationships

## 4.7 Outlier Detection

## 4.8 Feature Detecting

## 4.9 Findings

---

# 5.0 Analysis of Results

---

# 6.0 Discussion

---

# 7.0 Conclusions and Recommendations

---

# Reference List

---

# Appendix

