---
title: "Assessment 3: Group Project"
team: "Team Echo"
session: "Hexa 2, 2024"
coursecode: "ZZSC9020"
authors:
  - "Chuang, Keith – z5449930"
  - "Gandhi, Rupesh – z5368767"
Melbin, Joseph – z5394849
CHIU, Yiu Tong – z5039191
John Student (z123456), "
  - "Jim Student2 (zID), "
  - "Jack Student3 (zID)."
date: "2024-04-20"

output: html_document

title: "A Data Science Approach to Forecast Electricity Consumption in Australia"
team: 22
session: Term 1, 2021
coursecode: MATH0000
author: 
  - "John Student (z123456), "
  - "Jim Student2 (zID), "
  - "Jack Student3 (zID)." 
date: "25/07/2020"
Acknowledgements: 
  - "By far the greatest thanks must go to my supervisor for the guidance, care and support they provided."
  - "Thanks must also go to Emily, Michelle, John and Alex who helped by proof-reading the document in the final stages of preparation."
  - "Although I have not lived with them for a number of years, my family also deserve many thanks for their encouragement. Thanks go to Robert Taggart for allowing his thesis style to be shamelessly copied."


---

# Abstract

---

# Contents

---

# 1.0 Introduction

---

# 2.0 Literature Review

## 2.1 Industry/Operational Context

### 2.1.1 Variables of Interest
#### 2.1.1.1 Temperature
#### 2.1.1.2 Population
#### 2.1.1.3 Industry Base
#### 2.1.1.4 Renewable Engery Prevalence

## 2.2 Research Topic

## 2.3 Statistical Models

### 2.3.1 Regression Analysis

#### 2.3.1.1 Linear Models 

#### 2.3.1.2 Generalised Models

#### 2.3.1.3 Logistics Models

#### 2.3.1.4 Poisson Models

#### 2.3.1.5 Log-Linear Models

#### 2.3.1.6 Model Assessment

#### 2.3.1.7 Polynomial Models

#### 2.3.1.8 Splines

#### 2.3.1.9 Generalised Additive Models

### 2.3.2 Multivariate Analysis

---

# 3.0 Material and Methods

# Packages

```{r}

library(readr)
library(here)
library(ggplot2)

```



---

# 4.0 Exploratory Data Analysis

## 4.1 Area of Research
Determine the efficacy of temperature as the independent variable in prediction 
of energy demand in the subsequent 24 hours on a half hourly basis for NSW 
utilizing statistical models i.e. regression. Explore the predictive improvement 
through the addition of independent variables. Do these models provide 
comparable predictive results to the supplied predictions and if so under what 
circumstances.

## 4.2 Collecting and Loading Data

### 4.2.1 Provided Data 

``` {r}

# Load Data
data_forecastdemand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/a/forecastdemand_vic.csv"))

data_temperature_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_vic.csv"))

data_demand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_vic.csv"))

data_forecastdemand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/b/forecastdemand_sa.csv"))

data_temperature_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_sa.csv"))

data_demand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_sa.csv"))


data_forecastdemand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/c/forecastdemand_qld.csv"))

data_temperature_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_qld.csv"))

data_demand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_qld.csv"))

```

### 4.2.2 Sourced Data

``` {r}

```


``` {r} 
# QLD Data
ggplot(data_demand_qld, aes(x = DATETIME, y = TOTALDEMAND)) +
  geom_point() +
  labs(x = "X-axis label", y = "Y-axis label", title = "Scatter Plot")
```

---

# 5.0 Analysis of Results

---

# 6.0 Discussion

---

# 7.0 Conclusions and Recommendations

---

# Reference List

---

# Appendix

