---
title: "Project"
author: "Team Echo: "
date: "2024-03-19"
output: html_document
---

# Packages

```{r}

library(readr)
library(here)
library(ggplot2)

```

---

# Data

```{r}

# Load Data
data_forecastdemand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/a/forecastdemand_vic.csv"))

data_temperature_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_vic.csv"))

data_demand_vic <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_vic.csv"))

data_forecastdemand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/b/forecastdemand_sa.csv"))

data_temperature_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_sa.csv"))

data_demand_sa <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_sa.csv"))


data_forecastdemand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/c/forecastdemand_qld.csv"))

data_temperature_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/temprature_qld.csv"))

data_demand_qld <- read_csv(here("/Users/josephrobertmelbin/Documents/GitHub/Project2/data/Australia/d/totaldemand_qld.csv"))

```

---

# Abstract

---

# Contents

---

# 1.0 Introduction

---

# 2.0 Literature Review

## 2.1 Industry/Operational Context

### 2.1.1 Variables of Interest
#### 2.1.1.1 Temperature
#### 2.1.1.2 Population
#### 2.1.1.3 Industry Base
#### 2.1.1.4 Renewable Engery Prevalence

## 2.2 Research Topic

## 2.3 Statistical Models

### 2.3.1 Regression Analysis

#### 2.3.1.1 Linear Models 

#### 2.3.1.2 Generalised Models

#### 2.3.1.3 Logistics Models

#### 2.3.1.4 Poisson Models

#### 2.3.1.5 Log-Linear Models

#### 2.3.1.6 Model Assessment

#### 2.3.1.7 Polynomial Models

#### 2.3.1.8 Splines

#### 2.3.1.9 Generalised Additive Models

### 2.3.2 Multivariate Analysis

---

# 3.0 Material and Methods

---

# 4.0 Exploratory Data Analysis

``` {r}
ggplot(data_demand_qld, aes(x = DATETIME, y = TOTALDEMAND)) +
  geom_point() +
  labs(x = "X-axis label", y = "Y-axis label", title = "Scatter Plot")


```

---

# 5.0 Analysis of Results

---

# 6.0 Discussion

---

# 7.0 Conclusions and Recommendations

---

# Reference List

---

# Appendix

